Draft_Application_Stochastic_Volatility

### 2021-07-18 Real Dataset SV.Rmd
commit

In the following sections, we will try to replicate some of the analyses which were previously conducted on synthetic data, generated by a basic stochastic volatility model; this time, the same model specification and filters will be used on real data. However, as we will see, real data require different strategies to evaluate the effectiveness of several Sequential Monte Carlo techniques. Intuitively, such differences are due to the fact that now the latent process driving the Hidden Markov Model is not observed, nor do we know exactly its law of motion. 
This part of the work is structured as follows: section \textit{fill with the number of the section} provides a review of recent research in the stochastic volatility literature, including extensions of the basic specification and the use of particle filters; section \textit{fill with the number of the section} describes the data which we employ and the transformations that were applied to them. In section \textit{fill with the number of the section} we briefly compare the series of filtered states that are obtained from different particle filters; moreover, several PFs are compared in terms of computational cost in approximating the exact filtering distribution implied by our SV model. Section \textit{fill with the number of the section} features a forecast comparison between two SV specifications and a model that assumes constant mean and variance for the financial returns, namely a forecasting exercise that would not be possible were it not for particle filtering techniques. In section \textit{fill with the number of the section} we conclude by employing a non-parametric measure of realised volatility as a proxy for the latent volatility process, allowing a RMSE and MAE comparison between filters which is similar to that of section \textit{fill with the number of the section} for synthetic data.

@@

\section{Data}

@@

\textbf{Bibliography}\

>>>>

In the following sections, we will try to replicate some of the analyses which were previously conducted on synthetic data, generated by a basic stochastic volatility model; this time, the same model specification and filters will be used on real data. However, as we will see, real data require different strategies to evaluate the effectiveness of several Sequential Monte Carlo techniques. Intuitively, such differences are due to the fact that now the latent process driving the Hidden Markov Model is not observed, nor do we know exactly its law of motion. 
This part of the analysis is structured as follows: section \textit{fill with the number of the section} provides a review of recent research in the stochastic volatility literature, including extensions of the basic specification and the use of particle filters; section \textit{fill with the number of the section} describes the data which we employ and the transformations that were applied to them. In section \textit{fill with the number of the section} we briefly compare the series of filtered states that are obtained from different particle filters; moreover, several PFs are compared in terms of computational cost in approximating the exact filtering distribution implied by our SV model. Section \textit{fill with the number of the section} features a forecast comparison between two SV specifications and a model that assumes constant mean and variance for the financial returns, namely a forecasting exercise that would not be possible were it not for particle filtering techniques. In section \textit{fill with the number of the section} we conclude by employing a non-parametric measure of realised volatility as a proxy for the latent volatility process, allowing a RMSE and MAE comparison between filters which is similar to that of section \textit{fill with the number of the section} for synthetic data.

@@

\section{Data Description}

@@

\section{Visual Analysis after Filtering}

In this section, we offer a preliminary comparison of the various filtering methods, based on a visual analysis of graphs representing, for each index, the volatility of the observations obtained through the application of the different filtering techniques explored in the theoretical analysis. Plots that are compared represent the sequence of percentage log returns for a given index (the observations, in gray) and the sequence of filtered volatilities calculated using filtered states (in red) obtained through a specific filtering technique indicated in the caption of the plot.\footnote{Such filtered volatilities are obtained as follows. For each $t$, let us denote by $\mu_t$ the mean of the generated particles; the filtered volatility is then defined as $e^{\mu_t}$. Note that the number of particles was set at 10000 for each filter.} \

First, a general point is that the sequence of percentage log returns are different across the three indices. The reasons why this holds can be associated with the index-construction method and the characteristics of the economy and industrial setting it represents (SEE SECTION X). In particular, as previously noticed, there are extensive similarities between the DOW JONES Index and the S&P500, which can be visualized when looking at the percentage of log returns in the corresponding plots, since both are describing the US Economy. However, the STOXX50E differs quite extensively, especially in the different sensitivity to the Covid-19 Crisis and its consequences,  representing the European Economy. Aside from this note, the considerations expressed below on the sequences "filtered volatilities"\footnote{Per ora nei grafici c'è scritto "Filtered States", verrà corretto.} for the various techniques hold almost uniformly for all the three indices. Therefore, in the next paragraph we will be referring to the set of plots pertaining to one specific index (which could be any of the three).\

Generally speaking, note that, the sequence of "filtered volatilities" obtained from any of the filtering techniques follows roughly similar paths across all the plots. Unsurprisingly, the filtered volatility appears wider in periods corresponding to wider fluctuations of the percentage log returns and lower in periods corresponding to smaller such fluctuations. For example, all sequences of filtered volatilities show peaks corresponding to the higher fluctuations due to the 2020 Covid-19 Crisis outbreak and the consequent increases in uncertainty for the later periods.\

When focusing the attention on specific filters, first we can discuss the effect of resampling. Indeed, in the plot referring to the Sequential Importance Sampling procedure, the sequence of filtered volatilities is not able to reproduce the peaks as in with other filters, but also is characterized by collapsing confidence intervals. This is because ::::::: \
Then, if we instead look at the filtered volatilities obtained with resampling, we can appreciate how the filtered sequences for the Basic,\footnote{By this name, we refer to a particle filter that features no adaptive resampling; rather, resampling is conducted at each step of the algorithm.} Bootstrap, Auxiliary and Optimal Kernel Guided particle filters give rather analogous results, with no differences appearing at a first visual inspection.\footnote{Except for the Auxiliary Particle filtered sequence of volatilities, showing a slightly higher peak corresponding to the greatest volatility fluctuation in 2020, if compared to other filters in this subset. This is still valid for all three indices.} \
However, the sequence of filtered volatilities obtained by applying the Liu and West filter is clearly different from the others. Indeed, it shows a "more detailed" sequence of states (in the sense of the sequence of filtered volatilities being less smooth) and also a higher peak for the wide fluctuation corresponding to the 2020 Covid Crisis outbreak. Therefore, we may say that, compared to other filtering techniques, it offers a better representation of the data. This is because ... \textit{Potremmo parlare di maggiore sensibilità alle oscillazioni dei prezzi, o simili}

\section{Bibliography}\
